Intelligent Reflecting Surface-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design

### 2020-11-06 main.tex
:construction: Revise manuscript

			\textit{Third}, we conduct numerical evaluations that demonstrate the benefit of the proposed architecture. It is confirmed that the observations for SWIPT carries on to IRS-aided SWIPT. For example, dedicated power waveform can boost the energy transmission efficiency for multi-carrier transmission, while TS and PS are preferred at low and high SNR, respectively. Results also confirmed that the IRS bring a significant channel amplification and R-E enhancement with no impact on the waveform preference and transceiving strategy. Different from the active amplify-and-forward relay with optimal location around the midpoint \cite{Li2017}, the passive IRS should be placed next to either the transmitter or the receiver, due to the product-distance path loss model. It implies that that equipping the AP with an IRS can effectively extend the operation range of SWIPT systems. For active beamforming, increasing the number of transmit antennas brings a linear growth in SNR and a quadratic boost in harvested DC power. For passive beamforming, increasing the number of reflect elements provides a quadratic growth in SNR and a quartic boost in harvested DC power. Moreover, the adaptive IRS design outperforms the non-adaptive (rate/energy-optimized) ones for broadband transmission, while the performance of all three strategies coincides with the ideal Frequency-Selective (FS) IRS for narrowband transmission.

@@

		\begin{subsection}{Composite Channel}
			At subband $n$, denote the AP-user direct channel as $\boldsymbol{h}_{D,n}^H \in \mathbb{C}^{1 \times M}$, AP-IRS incident channel as $\boldsymbol{H}_{I,n} \in \mathbb{C}^{L \times M}$, and IRS-user reflective channel as $\boldsymbol{h}_{R,n}^H \in \mathbb{C}^{1 \times L}$. At the IRS, element $l$ ($l=1,\dots,L$) redistributes the incoming signal by adjusting the reflection coefficient $\phi_l=\alpha_l e^{j\theta_l}$ with reflection amplitude $\alpha_l \in [0,1]$ and phase shift $\theta_l \in [0,2\pi)$~\footnote{To investigate the performance upper bound of IRS, we suppose the reflection coefficient is maximized $\alpha_l=1\,,\forall l$ while the phase shift is a continuous variable over $[0,2\pi)$.}. Define the IRS matrix as $\boldsymbol{\Theta} = \mathrm{diag}(\phi_1, \dots, \phi_L) \in \mathbb{C}^{L \times L}$ that collects the reflection coefficients onto its main diagonal entries. The extra link introduced by IRS can be modeled as a concatenation of the AP-IRS incident channel, IRS reflection matrix, and IRS-user reflective channel. On top of this, the total composite channel is obtained by superposing the IRS-aided extra channel to the AP-user direct channel as
			\begin{equation}\label{eq:h_n}
				\boldsymbol{h}_{n}^H = \boldsymbol{h}_{D,n}^H + \boldsymbol{h}_{R,n}^H \boldsymbol{\Theta} \boldsymbol{H}_{I,n} = \boldsymbol{h}_{D,n}^H + \boldsymbol{\phi}^H \boldsymbol{V}_{n}
			\end{equation}
			where $\boldsymbol{\phi}=[\phi_1, \dots, \phi_L]^H \in \mathbb{C}^{L \times 1}$ and $\boldsymbol{V}_{n}=\mathrm{diag}(\boldsymbol{h}_{R,n}^H)\boldsymbol{H}_{I,n} \in \mathbb{C}^{L \times M}$. Note the conjugate transpose in the notation of $\boldsymbol{\phi}$ makes its entries the complex conjugate of the diagonal entries of $\boldsymbol{\Theta}$.
			\begin{remark}
				The IRS reflection coefficients are constant across the frequency, i.e. the IRS is frequency-flat where independent phase shift control over subbands are prohibited due to hardware constraints.
			\end{remark}
		\end{subsection}

@@

	\begin{section}{Problem Formulation}\label{se:problem_formulation}
		We characterize the R-E region through a current maximization problem subject to transmit power, IRS magnitude, and rate constraints
		\begin{maxi!}
			{\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho}{z(\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho)}{\label{op:original}}{\label{ob:original}}
			\addConstraint{\frac{1}{2}\left(\lVert{\boldsymbol{w}_I}\rVert^2+\lVert{\boldsymbol{w}_P}\rVert^2\right)\le{P}}\label{co:original_power}
			\addConstraint{R(\boldsymbol{\phi},\boldsymbol{w}_I,\rho) \ge \bar{R}}\label{co:original_rate}
			\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}\label{co:original_modulus}
			\addConstraint{0 \le \rho \le 1\,.}
		\end{maxi!}
		Problem~\eqref{op:original} is intricate as the variables are coupled and the objective function \eqref{ob:original}, the rate constraint \eqref{co:original_rate}, and the IRS magnitude constraint \eqref{co:original_modulus} are non-convex. To achieve a feasible solution, we propose a AO algorithm that iteratively updates the waveform and active beamforming at the transmitter, the phase shift at the IRS, and the power splitting ratio at the receiver, until convergence.
		\begin{remark}
			At subband $n$, the MIMO incident channel $\boldsymbol{H}_{I,n}$ can be divided into $L$ MISO channels between all transmit antennas and each IRS element. Reflection coefficient $\phi_l$ is shared for the $l$-th MISO incoming channel of all subbands, namely $MN$ entries in total. As information and power transfer favor different channel strength distribution over frequencies, for a given channel, the passive beamforming need to be optimized for each R-E point to maximize the R-E tradeoff. If the IRS is frequency-selective and there is only one transmit antenna, then reflector $l$ can simply align the AP-IRS($l$)-user channel with the AP-user channel at all subbands so that the joint design of waveform, active and passive beamforming, and splitting ratio would follow straightforwardly from existing SWIPT work \cite{Clerckx2018b}.
		\end{remark}

@@

			\begin{proposition}\label{pr:ao}
				Suppose $\boldsymbol{\Phi}^{\star}$ is rank-\num{1}, every limit point $(\boldsymbol{\phi}^{\star},\boldsymbol{w}_I^{\star},\boldsymbol{w}_P^{\star},\rho^{\star})$ of the proposed alternating algorithm is a stationary point of the original problem~\eqref{op:original}.
			\end{proposition}

			\begin{proof}\label{pf:ao}
				The objective function \eqref{ob:original} is non-decreasing over iterations of Algorithm~\ref{al:alternating}, which is also upper-bounded due to the unit-modulus constraint \eqref{co:original_modulus} and the average transmit power constraint \eqref{co:original_power}. Thus, Algorithm~\ref{al:alternating} is guaranteed to converge, namely the sequence $\{\boldsymbol{\phi}^{(i)},\boldsymbol{w}_I^{(i)},\boldsymbol{w}_P^{(i)},\rho^{(i)}\}$ generated by optimizing $\boldsymbol{\phi}$ and $\boldsymbol{w}_I,\boldsymbol{w}_P,\rho$ alternatively has limit points. As demonstrated in \cite{Grippo2000}, the solution is a stationary point of problem~\eqref{op:original}.
			\end{proof}
		\end{subsection}
	\end{section}

	\begin{section}{Performance Evaluations}\label{se:performance_evaluation}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{\columnwidth}
			\import{assets/}{layout.pdf_tex}
			\caption{System layout in the simulation.}
			\label{fi:layout}
		\end{figure}
		To evaluate the performance of the proposed IRS-aided SWIPT system, we characterize the average R-E regions under typical setups. Consider a large open space WiFi-like environment at a center frequency of \SI{5.18}{\GHz} with reference bandwidth $B=1\,\si{\MHz}$. As shown in Fig.~\ref{fi:layout}, we assume the IRS moves along a horizontal line parallel to the AP-user path and let $d_H$, $d_V$ be the horizontal and vertical distances from the AP to the IRS, respectively. We also denote $d_D$, $d_I$, $d_R$ as the length of direct, incident, reflective paths such that $d_I=\sqrt{d_H^2+d_V^2}$, $d_R=\sqrt{(d_D-d_H)^2+d_V^2}$. We choose as reference $d_D=15\,\si{\meter}$, $d_V=2\,\si{\meter}$ and $d_H=2\,\si{\meter}$. The path loss and fading parameters are obtained from IEEE TGn channel model D \cite{Erceg2004}, and reference path loss is set to $L_0=-35\,\si{\dB}$ at $d_0=1\,\si{\meter}$. We assume all channels are NLoS with taps modelled as i.i.d. CSCG random variables of unit average sum-power to create a normalized multipath response. The reference numbers of transmit antennas, IRS reflectors and subbands are $M=1$, $L=20$, $N=16$, respectively. No spatial correlation is considered across transmit antennas and IRS arrays. Rectenna parameters are taken as $k_2=0.0034$, $k_4=0.3829$, $R_{\text{ant}}=50\,\si{\ohm}$. With \SI{0}{\dBi} transmit antenna gain, the average Effective Isotropic Radiated Power (EIRP) is fixed to $P=-36\,\si{\dBm}$ while the reference average noise power is $\sigma_n=-40\,\si{\dBm}$ at all subbands. We also assume \SI{0}{\dBi} IRS element gain and \SI{2}{\dBi} receive antenna gain. For the algorithm, the tolerance is $\epsilon=10^{-8}$, the number of candidates in the Gaussian randomization method is $Q=10^{3}$, and the R-E region is averaged over \num{200} channel realizations. In the R-E boundary, the leftmost point corresponds to WPT ($\rho=1$) where power can be allocated simultaneously to modulated and unmodulated waveform to maximize the average output DC current.
		% - at the WPT point, power is allocated to the multisine waveform only, isn't it?
		% - for single-carrier WPT, modulated waveform may be preferred as it has a modulation gain (also proved in Fig. 5)
		On the other hand, the rightmost point corresponds to WIT ($\rho=0$) where the solution coincides with the Water-Filling (WF) algorithm that allocates all power to modulated waveform only. For a fair comparison, the $x$-axis of the plots has been normalized to per-subband rate.

@@

		Fig.~\ref{fi:re_subband} illustrates the average R-E region versus the number of subband $N$. \textit{First}, it is observed that increasing $N$ reduces the per-subband rate but boosts the harvested energy. The reason is that although each subband receives a smaller proportion of the total power, more balanced terms $m_P$ are introduced to further amplify the output DC current, as suggested by the scaling laws in \cite{Clerckx2018b}. Sorted waveform amplitudes in Fig.~\ref{fi:waveform_subband} also confirmed that from the perspective of WPT, dedicated multisine waveform is unnecessary for a small $N$ but is required for a large $N$. As shown in \eqref{eq:y_I4} and \eqref{eq:y_P4}, the only difference of modulated and unmodulated waveform on $z$ exists in the fourth-order terms, where $\mathcal{E}\left\{\mathcal{A}\left\{y_{I}^4(t)\right\}\right\}$ has $N^2$ monomials with a modulation gain of \num{2} and $\mathcal{A}\left\{y_{P}^4(t)\right\}$ has $(2N^3+N)/3$ monomials without modulation gain. Therefore, superposed waveform enlarges the R-E region for a sufficiently large $N$ (typically no smaller than 4). However, an excessively large $N$ not only increases complexity but also operates out of the small-signal harvester model, thus become prohibitive. \textit{Second}, the R-E region is convex for $N = 2, 4$ and concave-convex for $N = 8, 16$. This has the consequence that PS outperforms TS for a small $N$ and is outperformed for a large $N$. When $N$ is in between, the optimal strategy is a combination of both, i.e. a time sharing between the WPT point and the tangent WIPT point obtained by PS. Compared with the linear harvester model that requires no dedicated power waveform and always prefer PS, the rectifier nonlinearity enlarges the R-E region by favouring a different waveform and transceiving strategy, both heavily depends on $N$.

@@

		The influence of the average noise power on the average R-E region is investigated in Fig.~\ref{fi:re_noise}. We \textit{first} note that for a large number of subbands, the R-E region is approximately concave for a high noise level and approximately convex for a low noise level. Hence, TS is preferred at low SNR while PS is preferred at high SNR. This is because at a low SNR, the capacity-achieving WF algorithm tends to allocate more power to few strongest subbands. As the rate constraint $\bar{R}$ decreases, more subbands are activated to further boost the harvested energy via the coupling effect by rectifier nonlinearity. \textit{Second}, there exists a turning point in the R-E region especially for a small noise ($\sigma_n \le -40\,\si{\dBm}$). The reason is that when $\bar{R}$ departs slightly from the maximum achievable rate, the algorithm mainly adjusts the splitting ratio $\rho$ rather than put more weight on the multisine waveform, as a small amplitude could be inefficient for energy maximization. On the other hand, as $\bar{R}$ further reduces, a modulated waveform with a very large $\rho$ could be outperformed by a superposed waveform with a smaller $\rho$, due to advantage the of multisine. The result highlights the benefit of multisine and the necessity of joint optimization of waveform and splitting ratio.

@@

		In Fig.~\ref{fi:re_distance}, we compare the average R-E region achieved by different AP-IRS horizontal distance $d_H$. A \textit{first} observation is that, different from the active amplify-and-forward relay that achieves optimal performance when developed around the midpoint, placing the IRS closer to either the AP or the user would further improve the R-E tradeoff. It origins from the product-distance path loss model that applies to finite-size element reflection. As shown in Fig.~\ref{fi:path_loss}, although the piecewise TGn path loss model further penalizes large distance (greater than \SI{10}{\meter} for model D), it is still beneficial to have a short-long or long-short transmission setup, since signal attenuation increases fast at a short distance and experiences marginal effect at a long distance. On the other hand, it also suggests that developing an IRS next to the AP can effectively extend the operation range of SWIPT systems. Considering the passive characteristic of IRS, opportunities are that it can be directly supported by the SWIPT network. A \textit{second} observation is that there exist two optimal IRS development locations that maximizes the path loss production $\Lambda_I\Lambda_R$. It implies that more than one IRS may be implemented to further enlarge the R-E region, one attached to the AP and one attached to the IRS.

@@

		The number of transmit antennas $M$ and IRS reflectors $L$ influence the average R-E tradeoff as revealed in \ref{fi:re_tx} and Fig.~\ref{fi:re_reflector}. \textit{First,} it is observed that adding either active or passive elements benefits both information and power transmission while preserving the concavity-convexity of the R-E region. This is because increasing $M$ or $L$ indeed enhances the equivalent composite channel strength such that only the magnitude of the terms in \eqref{eq:z_expand} is amplified. Therefore, it is concluded that the number of active and passive elements have no impact on the waveform preference and transceiving strategy. \textit{Second,} the R-E benefit of passive beamforming scales faster than that of active beamforming. This behavior is more obvious in Fig.~\ref{fi:scaling_tx} and \ref{fi:scaling_reflector} that illustrate the performance of WIT and WPT versus $M$ and $L$. For the active MRT beamforming, doubling $M$ brings a \SI{3}{\dB} gain at the average output SNR, corresponding to a transmit array gain of $M$ and a doubled harvester input power. Due to the rectifier nonlinearity, the output DC current ends up with a nearly four-time (\SI{12}{\dB}) increase, validating the active scaling law in the order of $M^2$ \cite{Clerckx2016a}. On the other hand, when the IRS is very closed to the transmitter or the receiver, doubling $L$ can bring a SNR boost up to \SI{6}{\dB}, corresponding to an reflect array gain of $L^2$. An interpretation is that the IRS collects $L$ signal copies with a receive array gain $L$, then performs an equal gain reflection with a transmit array gain $L$, achieving a total gain of $L^2$. As a consequence, doubling $L$ brings a four-fold boost on the received signal power that further amplifies the harvested DC current by 16 times (\SI{24}{\dB}), suggesting a passive scaling law in the order of $L^4$. Besides, by considering harvester nonlinearity, the proposed design brings a \SI{20}{\dB} increase in the output DC current over conventional linear perspective. However, the diode nonlinear model is based on a small-signal expansion and the input-output relationship would be linear in the high-power regime. These results emphasize the effectiveness of passive beamforming and the importance of modelling rectifier nonlinearity in the system design. \textit{Third,} using only few transmit antenna can achieve the performance of numerous reflectors in a much lower complexity. The reason is that the system performance in our setup is dominated by the AP-user direct link, with the optimal active beamformer obtained in closed form. As shown in Fig.~\ref{fi:path_loss}, the direct path loss $\Lambda_D$ is in the scope of \num{e-7} while the extra path loss product $\Lambda_I\Lambda_R$ is below \num{e-10}. Therefore, although increasing $L$ can significantly enhance the AP-IRS-user extra channel, its amplitude is still too small compared with the direct channel such that increasing $M$ can be more efficient for a practical configuration. Another explanation is that the RF-chains of active antennas provide design flexibility in the frequency domain while the frequency-flat characteristic of IRS restricts its benefit for broadband transmission. On the contrary, a large number of active antennas suffer from channel hardening while the IRS can use adaptive reflection to potentially mitigate the effect. As a consequence, active beamforming is more effective for small-scale SWIPT where the direct channel dominates the transmission, and passive beamforming can be more effective for large-scale SWIPT where the extra channel dominates the transmission.

		\begin{figure}[!t]
			\centering
			\begin{adjustbox}{width=\linewidth}
				\input{assets/re_irs_1mhz.tex}
			\end{adjustbox}
			\caption{Average R-E region for adaptive, ideal, fixed and no IRS over $B=1\,\si{\MHz}$ for $M=1$, $N=16$, $L=20$, $\sigma_n=-40\,\si{\dBm}$ and $d_H=d_V=2\,\si{\meter}$.}
			\label{fi:re_irs_1mhz}
		\end{figure}

>>>>

			\textit{Third}, we conduct numerical evaluations that demonstrate the benefit of the proposed architecture. It is confirmed that the observations for SWIPT carries on to IRS-aided SWIPT. For example, dedicated power waveform can boost the energy transmission efficiency for multi-carrier transmission, while TS and PS are preferred at low and high SNR, respectively. Results also confirmed that the IRS bring a significant channel amplification and R-E enhancement with no impact on the waveform preference and transceiving strategy. Different from the active amplify-and-forward (AF) relay with optimal location around the midpoint \cite{Li2017}, the passive IRS should be placed next to either the transmitter or the receiver, due to the product-distance path loss model. It implies that that equipping the AP with an IRS can effectively extend the operation range of SWIPT systems. For active beamforming, increasing the number of transmit antennas brings a linear growth in SNR and a quadratic boost in harvested DC power. For passive beamforming, increasing the number of reflect elements provides a quadratic growth in SNR and a quartic boost in harvested DC power. Moreover, the adaptive IRS design outperforms the non-adaptive (rate/energy-optimized) ones for broadband transmission, while the performance of all three strategies coincides with the ideal Frequency-Selective (FS) IRS for narrowband transmission.

@@

		\begin{subsection}{Composite Channel}
			At subband $n$, denote the AP-user direct channel as $\boldsymbol{h}_{D,n}^H \in \mathbb{C}^{1 \times M}$, AP-IRS incident channel as $\boldsymbol{H}_{I,n} \in \mathbb{C}^{L \times M}$, and IRS-user reflective channel as $\boldsymbol{h}_{R,n}^H \in \mathbb{C}^{1 \times L}$. At the IRS, element $l$ ($l=1,\dots,L$) redistributes the incoming signal by adjusting the reflection coefficient $\phi_l=\alpha_l e^{j\theta_l}$ with reflection amplitude $\alpha_l \in [0,1]$ and phase shift $\theta_l \in [0,2\pi)$~\footnote{To investigate the performance upper bound of IRS, we suppose the reflection coefficient is maximized $\alpha_l=1\,,\forall l$ while the phase shift is a continuous variable over $[0,2\pi)$.}. Define the IRS matrix as $\boldsymbol{\Theta} = \mathrm{diag}(\phi_1, \dots, \phi_L) \in \mathbb{C}^{L \times L}$ that collects the reflection coefficients onto its main diagonal entries. The extra link introduced by IRS can be modeled as a concatenation of the AP-IRS incident channel, IRS reflection matrix, and IRS-user reflective channel. On top of this, the total composite channel is obtained by superposing the IRS-aided auxiliary channel to the AP-user direct channel as
			\begin{equation}\label{eq:h_n}
				\boldsymbol{h}_{n}^H = \boldsymbol{h}_{D,n}^H + \boldsymbol{h}_{R,n}^H \boldsymbol{\Theta} \boldsymbol{H}_{I,n} = \boldsymbol{h}_{D,n}^H + \boldsymbol{\phi}^H \boldsymbol{V}_{n}
			\end{equation}
			where $\boldsymbol{\phi}=[\phi_1, \dots, \phi_L]^H \in \mathbb{C}^{L \times 1}$ and $\boldsymbol{V}_{n}=\mathrm{diag}(\boldsymbol{h}_{R,n}^H)\boldsymbol{H}_{I,n} \in \mathbb{C}^{L \times M}$. Note the conjugate transpose in the notation of $\boldsymbol{\phi}$ makes its entries the complex conjugate of the diagonal entries of $\boldsymbol{\Theta}$.
			\begin{remark}\label{re:irs_frequency_flat}
				Due to hardware constraints, independent phase shift control for different frequencies is prohibited and the IRS is frequency-flat. Hence, the auxiliary channels at different frequencies cannot be simultaneously maximized and there is a tradeoff in subchannel alignment.
				% At subband $n$, the AP-IRS MIMO channel $\boldsymbol{H}_{I,n}$ can be divided into $L$ MISO channels between all transmit antennas and each IRS element. Reflection coefficient $\phi_l$ is shared for the $l$-th AP-IRS MISO channel of all subbands, namely $MN$ entries in total.
				The reflection coefficient of each IRS element is shared for the AP-IRS MISO channel and over all subbands, namely $MN$ entries in total.
			\end{remark}
		\end{subsection}

@@

	\begin{section}{Problem Formulation}\label{se:problem_formulation}
		We characterize the R-E region through a current maximization problem subject to transmit power, IRS magnitude, and rate constraints
		\begin{maxi!}
			{\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho}{z(\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho)}{\label{op:original}}{\label{ob:original}}
			\addConstraint{\frac{1}{2}\left(\lVert{\boldsymbol{w}_I}\rVert^2+\lVert{\boldsymbol{w}_P}\rVert^2\right)\le{P}}\label{co:original_power}
			\addConstraint{R(\boldsymbol{\phi},\boldsymbol{w}_I,\rho) \ge \bar{R}}\label{co:original_rate}
			\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}\label{co:original_modulus}
			\addConstraint{0 \le \rho \le 1\,.}
		\end{maxi!}
		Problem~\eqref{op:original} is intricate as the variables are coupled and the objective function \eqref{ob:original}, the rate constraint \eqref{co:original_rate}, and the IRS magnitude constraint \eqref{co:original_modulus} are non-convex. To achieve a feasible solution, we propose a AO algorithm that iteratively updates the waveform and active beamforming at the transmitter, the phase shift at the IRS, and the power splitting ratio at the receiver, until convergence.
		\begin{remark}\label{re:irs_subchannel_alignment}
			As information and power transfer prefer different subchannel strength distribution, the frequency-flat characteristic of IRS introduces a resource allocation problem. Following Remark \ref{re:irs_frequency_flat}, if the IRS is frequency-selective and there is only one transmit antenna, then each reflector can simultaneously align the corresponding AP-IRS-user channel with the AP-user channel at all subbands. In this case, the strength of all subchannels are maximized so that the joint design would follow straightforwardly from existing SWIPT work \cite{Clerckx2018b}.
		\end{remark}

@@

			\begin{proposition}\label{pr:ao}
				Suppose $\boldsymbol{\Phi}$ is rank-\num{1}, every limit point $(\boldsymbol{\phi}^{\star},\boldsymbol{w}_I^{\star},\boldsymbol{w}_P^{\star},\rho^{\star})$ of the proposed alternating algorithm is a stationary point of the original problem~\eqref{op:original}.
			\end{proposition}

			\begin{proof}\label{pf:ao}
				The objective function \eqref{ob:original} is non-decreasing over iterations of Algorithm~\ref{al:alternating}, which is also upper-bounded due to the unit-modulus constraint \eqref{co:original_modulus} and the average transmit power constraint \eqref{co:original_power}. If $\boldsymbol{\Phi}$ is rank-\num{1}, then $\boldsymbol{\phi}$ can be obtained by EVD without performance loss, and the sequence $\{\boldsymbol{\phi}^{(i)},\boldsymbol{w}_I^{(i)},\boldsymbol{w}_P^{(i)},\rho^{(i)}\}$ generated by alternatively optimizing $\boldsymbol{\phi}$ and $\boldsymbol{w}_I,\boldsymbol{w}_P,\rho$ has limit points so that Algorithm~\ref{al:alternating} is guaranteed to converge. As demonstrated in \cite{Grippo2000}, the solution is a stationary point of problem~\eqref{op:original}.
			\end{proof}
		\end{subsection}
	\end{section}

	\begin{section}{Performance Evaluations}\label{se:performance_evaluation}
		\begin{figure}[!t]
			\centering
			\def\svgwidth{\columnwidth}
			\import{assets/}{layout.pdf_tex}
			\caption{System layout in the simulation.}
			\label{fi:layout}
		\end{figure}
		To evaluate the performance of the proposed IRS-aided SWIPT system, we characterize the average R-E regions under typical setups. Consider a large open space WiFi-like environment at a center frequency of \SI{5.18}{\GHz} with reference bandwidth $B=1\,\si{\MHz}$. As shown in Fig.~\ref{fi:layout}, we assume the IRS moves along a horizontal line parallel to the AP-user path and let $d_H$, $d_V$ be the horizontal and vertical distances from the AP to the IRS, respectively. We also denote $d_D$, $d_I$, $d_R$ as the length of direct, incident, reflective paths such that $d_I=\sqrt{d_H^2+d_V^2}$, $d_R=\sqrt{(d_D-d_H)^2+d_V^2}$. We choose as reference $d_D=15\,\si{\meter}$, $d_V=2\,\si{\meter}$ and $d_H=2\,\si{\meter}$. The path loss and fading parameters are obtained from IEEE TGn channel model D \cite{Erceg2004}, and reference path loss is set to $L_0=-35\,\si{\dB}$ at $d_0=1\,\si{\meter}$. We assume all channels are NLoS with taps modelled as i.i.d. CSCG random variables of unit average sum-power to create a normalized multipath response. The reference numbers of transmit antennas, IRS reflectors and subbands are $M=1$, $L=20$, $N=16$, respectively. No spatial correlation is considered across transmit antennas and IRS arrays. Rectenna parameters are taken as $k_2=0.0034$, $k_4=0.3829$, $R_{\text{ant}}=50\,\si{\ohm}$. With \SI{0}{\dBi} transmit antenna gain, the average Effective Isotropic Radiated Power (EIRP) is fixed to $P=-36\,\si{\dBm}$ while the reference average noise power is $\sigma_n=-40\,\si{\dBm}$ at all subbands. We also assume \SI{0}{\dBi} IRS element gain and \SI{2}{\dBi} receive antenna gain. For the algorithm, the tolerance is $\epsilon=10^{-8}$, the number of candidates in the Gaussian randomization method is $Q=10^{3}$, and the R-E region is averaged over \num{200} channel realizations. In the R-E boundary, the leftmost point corresponds to WPT ($\rho=1$) where power can be allocated simultaneously to modulated and unmodulated waveform to maximize the average output DC current.
		% - at the WPT point, power is allocated to the multisine waveform only, isn't it?
		% - for single-carrier WPT, modulated waveform may be preferred as it has a modulation gain (also proved in Fig. 5)
		On the other hand, the rightmost point corresponds to WIT ($\rho=0$) where the solution coincides with the Water-Filling (WF) algorithm that allocates all power to modulated waveform only. For a fair comparison, the $x$-axis of the plots has been normalized to average subband rate $R/N$.

@@

		Fig.~\ref{fi:re_subband} illustrates the average R-E region versus the number of subband $N$. \textit{First}, it is observed that increasing $N$ reduces the average per-subband rate $R/N$ but boosts the harvested energy. The reason is that the power budget is divided into smaller portions for each subband, but more balanced terms $m_P$ are introduced to further amplify the output DC current, as suggested by the scaling laws in \cite{Clerckx2018b}. Sorted waveform amplitudes in Fig.~\ref{fi:waveform_subband} also confirmed that from the perspective of WPT, dedicated multisine waveform is unnecessary for a small $N$ but is required for a large $N$. As shown in \eqref{eq:y_I4} and \eqref{eq:y_P4}, the only difference between modulated and unmodulated waveform on $z$ exists in the fourth-order terms, where $\mathcal{E}\left\{\mathcal{A}\left\{y_{I}^4(t)\right\}\right\}$ has $N^2$ monomials with a modulation gain of \num{2} and $\mathcal{A}\left\{y_{P}^4(t)\right\}$ has $(2N^3+N)/3$ monomials without modulation gain. Therefore, superposed waveform enlarges the R-E region for a sufficiently large $N$ (typically no smaller than 4). \textit{Second}, the R-E region is convex for $N = 2, 4$ and concave-convex for $N = 8, 16$. This has the consequence that PS outperforms TS for a small $N$ and is outperformed for a large $N$. When $N$ is in between, the optimal strategy is a combination of both, i.e. a time sharing between the WPT point and the tangent WIPT point obtained by PS. Compared with the linear harvester model that requires no dedicated power waveform and always prefer PS, the rectifier nonlinearity enlarges the R-E region by favouring a different waveform and transceiving strategy, both heavily depends on $N$.

@@

		The influence of the average noise power on the average R-E region is investigated in Fig.~\ref{fi:re_noise}. \textit{First}, we note that for a large number of subbands, the R-E region is approximately concave for a high noise level and approximately convex for a low noise level. Hence, TS is preferred at low SNR while PS is preferred at high SNR. This is because at a low SNR, the capacity-achieving WF algorithm tends to allocate more power to few strongest subbands. As the rate constraint $\bar{R}$ decreases, more subbands are activated to further boost the harvested energy via the coupling effect by rectifier nonlinearity. \textit{Second}, there exists a turning point in the R-E region especially for a small noise ($\sigma_n \le -40\,\si{\dBm}$). The reason is that when $\bar{R}$ departs slightly from the maximum achievable rate, the algorithm mainly adjusts the splitting ratio $\rho$ rather than put more weight on the multisine waveform, as a small amplitude could be inefficient for energy maximization. On the other hand, as $\bar{R}$ further reduces, a modulated waveform with a very large $\rho$ could be outperformed by a superposed waveform with a smaller $\rho$, due to the advantage the of multisine for efficient power transfer. The result highlights the benefit of multisine and the necessity of joint optimization of waveform and splitting ratio.

@@

		In Fig.~\ref{fi:re_distance}, we compare the average R-E region achieved by different AP-IRS horizontal distance $d_H$. A \textit{first} observation is that, different from the active AF relay that achieves optimal performance when developed around the midpoint, placing the IRS closer to either the AP or the user would further improve the R-E tradeoff. It origins from the product-distance path loss model that applies to finite-size element reflection. As shown in Fig.~\ref{fi:path_loss}, although the piecewise TGn path loss model further penalizes large distance (greater than \SI{10}{\meter} for model D), it is still beneficial to have a short-long or long-short transmission setup, since signal attenuation increases fast at a short distance and experiences marginal effect at a long distance. On the other hand, it also suggests that developing an IRS next to the AP can effectively extend the operation range of SWIPT systems. Considering the passive characteristic of IRS, opportunities are that it can be directly supported by the SWIPT network. A \textit{second} observation is that there exist two optimal IRS development locations that minimizes the path loss production $\Lambda_I\Lambda_R$. It implies that more than one IRS may be implemented to further enlarge the R-E region, one attached to the AP and one attached to the IRS.

@@

		The number of transmit antennas $M$ and IRS reflectors $L$ influence the average R-E tradeoff as revealed in \ref{fi:re_tx} and Fig.~\ref{fi:re_reflector}. \textit{First,} it is observed that adding either active or passive elements benefits both information and power transmission while preserving the concavity-convexity of the R-E region. This is because increasing $M$ or $L$ indeed enhances the equivalent composite channel strength such that only the magnitude of the terms in \eqref{eq:z_expand} is amplified. Therefore, it is concluded that the number of active and passive elements have no impact on the waveform preference and transceiving strategy. \textit{Second,} the R-E benefit of passive beamforming scales faster than that of active beamforming. This behavior is more obvious in Fig.~\ref{fi:scaling_tx} and \ref{fi:scaling_reflector} that illustrate the performance of WIT and WPT versus $M$ and $L$. For the active MRT beamforming, doubling $M$ brings a \SI{3}{\dB} gain at the average output SNR, corresponding to a transmit array gain of $M$ and a doubled harvester input power. Due to the rectifier nonlinearity, the output DC current ends up with a nearly four-time (\SI{12}{\dB}) increase, validating the active scaling law in the order of $M^2$ \cite{Clerckx2016a}. On the other hand, when the IRS is very close to the transmitter or the receiver, doubling $L$ can bring a SNR boost up to \SI{6}{\dB}, corresponding to a reflect array gain of $L^2$. An interpretation is that the IRS collects $L$ signal copies with a receive array gain $L$, then performs an equal gain reflection with a transmit array gain $L$, achieving a total gain of $L^2$. As a consequence, doubling $L$ brings a four-fold boost on the received signal power that further amplifies the harvested DC current by 16 times (\SI{24}{\dB}), suggesting a passive scaling law in the order of $L^4$. Besides, by considering harvester nonlinearity, the proposed design brings a \SI{20}{\dB} increase in the output DC current over conventional linear perspective. These results emphasize the effectiveness of passive beamforming and the importance of modelling rectifier nonlinearity in the system design \footnote{However, the diode nonlinear model is based on a small-signal expansion and the input-output relationship would be linear in the high-power regime, which prevents an excessively large $M$ or $L$.}. \textit{Third,} using only few transmit antenna can achieve the performance of numerous reflectors in a much lower complexity. The reason is that the system performance in our setup is dominated by the AP-user direct link, with the optimal active beamformer obtained in closed form. As shown in Fig.~\ref{fi:path_loss}, the direct path loss $\Lambda_D$ is in the scope of \num{e-7} while the extra path loss product $\Lambda_I\Lambda_R$ is below \num{e-10}. Therefore, although increasing $L$ can significantly enhance the AP-IRS-user auxiliary channel, its amplitude is still too small compared with the direct channel such that increasing $M$ can be more efficient for a practical configuration. Another explanation is that the RF-chains of active antennas provide design flexibility in the frequency domain while the frequency-flat characteristic of IRS restricts its benefit for broadband transmission. On the contrary, a large number of active antennas suffer from channel hardening while the IRS can use adaptive reflection to potentially mitigate the effect. As a consequence, active beamforming is more effective for small-scale SWIPT where the direct channel dominates the transmission, and passive beamforming can be more effective for large-scale SWIPT where the auxiliary channel dominates the transmission.

		\begin{figure}[!t]
			\centering
			\begin{adjustbox}{width=\linewidth}
				\input{assets/re_irs_1mhz.tex}
			\end{adjustbox}
			\caption{Average R-E region for ideal, adaptive, fixed and no IRS over $B=1\,\si{\MHz}$ for $M=1$, $N=16$, $L=20$, $\sigma_n=-40\,\si{\dBm}$ and $d_H=d_V=2\,\si{\meter}$.}
			\label{fi:re_irs_1mhz}
		\end{figure}

		\begin{figure}[!t]
			\centering
			\begin{adjustbox}{width=\linewidth}
				\input{assets/re_irs_10mhz.tex}
			\end{adjustbox}
			\caption{Average R-E region for ideal, adaptive, fixed and no IRS over $B=10\,\si{\MHz}$ for $M=1$, $N=16$, $L=20$, $\sigma_n=-40\,\si{\dBm}$ and $d_H=d_V=2\,\si{\meter}$.}
			\label{fi:re_irs_10mhz}
		\end{figure}
